---
title: "Instrument comparison report"
author: "Haifa Splittorff"

output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
    includes:
      in_header: olink_style_windows.tex
      before_body: olink_title_style_internal_projects.tex
  html_document:
    df_print: paged
fontsize: 12pt
---
\makeolinktitle
\thispagestyle{olink_header_toc}
\tableofcontents
\newpage
\pagestyle{olink_header}


```{r setup, include=FALSE}
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(fig.pos = 'H',tab.pos = "H", out.extra = "")
knitr::opts_knit$set(eval.after = "fig.cap")

source("mylib.R")
sapply(list.files("Functions", pattern ="\\.R$", full.names = TRUE), source)

loadfonts(device = "win")

windowsFonts("Arial" = windowsFont("Arial"))
par(family="Arial")
plot(1:10)

options(knitr.table.format = "pandoc")


# Load configuration
config <- yaml::read_yaml("config_instrument.yml")

input_data <- config$input_data
parameters <- config$parameters
plot_settings <- config$plot_settings


# Output directory
product_name <- parameters$Product
project_name <- parameters$Project
output_dir <- file.path("Results", product_name, "Instrument_comparison")
yourdata <- readRDS(input_data)
matrix_types <- unique(yourdata[[as.character(sym(parameters$matrix_column))]] )



```

\newpage
\section{Param}

Product :  `r product_name`   
Project: `r project_name`   

\newpage

\section{PCA}


```{r, echo=FALSE, results='asis'}

  cat("![PCA plot for samples using NPX data](", output_dir, "/PCA_plot.", plot_settings$format, "){width=100%}",sep="")

```
\newpage


\section{IQR}


```{r, echo=FALSE, results='asis'}

  cat("![IQR plot for samples using NPX data](", output_dir, "/qc_plot.", plot_settings$format, "){width=100%}",sep="")

```

\newpage


\section{Sample wise NPX distribution}


```{r, echo=FALSE, results='asis'}

  cat("![Sample-wise NPX distribution](", output_dir, "/distplot.", plot_settings$format, "){width=100%}",sep="")

```
\newpage


\section{QC summary}

```{r}

generate_qc_table(yourdata, compare_by = parameters$run_var , qc_col = "sample_qc",
                              caption = "QC Status of Samples", font_size = 12, format = "latex")

```



```{r}

generate_qc_table(yourdata, compare_by =parameters$run_var, qc_var = "assay", qc_col = "assay_qc",
                              caption = "QC Status of assays", font_size = 12, format = "latex")

```


\newpage

\section{Total number of counts}

```{r}
generate_count_table_with_block(yourdata, comp_var=parameters$run_var,
                                              values_fill = 0, font_size = 12, format = "latex")
```


\newpage
\section{Count distribution within categories}

```{r}

result<- descriptive_tabulation(yourdata,parameters$run_var)

result$count_categ_table

```




\section{Count distribution}



```{r, echo=FALSE, results='asis'}

  cat("![Count distribution](", output_dir, "/count_density.", plot_settings$format, "){width=80%}",sep="")

```




\section{NPX distribution}


```{r, echo=FALSE, results='asis'}

  cat("![NPX distribution](", output_dir, "/NPX_density.", plot_settings$format, "){width=80%}",sep="")

```




\newpage


\section{Count correlations}


\section{Per assay count correlations}

```{r, corcount}
result <- correlation_function(
  data = filter(yourdata, if (parameters$Background == "detection") {
        get(parameters$Background) == "above EOB"  
      } else {
      NPX > get(parameters$Background) }),
  value_column = "log2count",
  comparison_column = parameters$run_var,
  filter_sample_type = "SAMPLE",
  filter_assay_type = "assay",
)

create_correlation_table(result$summary , font_size = 8, format = "latex")


```


\section{Per assay NPX correlations}


```{r, NPXcor}

result <- correlation_function(
  data = filter(yourdata, if (parameters$Background == "detection") {
        get(parameters$Background) == "above EOB"  
      } else {
      NPX > get(parameters$Background) }),
  value_column = "NPX",
  comparison_column = parameters$run_var,
  filter_sample_type = "SAMPLE",
  filter_assay_type = "assay",
)

create_correlation_table(result$summary , font_size = 8, format = "latex")

```

\newpage


\section{NPX variance differences across instruments}

```{r}
var<- variance_function(yourdata, parameters$run_var)

var$summary %>%
  kable(caption = "\\label{tab:var} Differences in variances across instruments.")

```



\newpage


\section{Count regression analysis}

```{r, results='asis', echo=FALSE}
rmd_syntax<- generate_rmd_syntax_for_figures(folder_path=paste0(output_dir,"/regression_plots_log2count_",parameters$run_var), label_prefix = "countreg", width = "100%")
cat(rmd_syntax, sep= "\n\n")
```




\newpage


\section{NPX regression analysis}

```{r, results='asis', echo=FALSE}
rmd_syntax<- generate_rmd_syntax_for_figures(folder_path=paste0(output_dir,"/regression_plots_NPX_",parameters$run_var), label_prefix = "npxreg", width = "100%")
cat(rmd_syntax, sep= "\n\n")
```

\newpage

\section{External controls}


```{r, echo=FALSE, results='asis'}

  cat("![Sample Control NPX distribution](", output_dir, "/distplot_SAMPLE_CONTROL.", plot_settings$format, "){width=100%}",sep="")

```

\newpage


```{r, echo=FALSE, results='asis'}

  cat("![Negative Control NPX distribution](", output_dir, "/distplot_NEGATIVE_CONTROL.", plot_settings$format, "){width=100%}",sep="")

```


\newpage


```{r, echo=FALSE, results='asis'}

  cat("![Plate Control NPX distribution](", output_dir, "/distplot_PLATE_CONTROL.", plot_settings$format, "){width=100%}",sep="")

```


\newpage
\section{Internal controls}


```{r, echo=FALSE, results='asis'}

  cat("![Internal Control log2count regression](", output_dir, "/Internal_control_regression.", plot_settings$format, "){width=100%}",sep="")

```

\newpage
\section{Coefficient of variance based on samples controls}


```{r}

cv<-generate_cv_summary(
    df = filter(yourdata,sample_qc=="PASS"),
    comp_var = parameters$run_var,
    npx = parameters$npx_var,
    threshold = parameters$Background,
    group_vars = c(parameters$run_var, "plate_id", "block"),
    pivot_var = parameters$run_var,
    value_col = "intra_CV",
    caption = "Average of intra plate CV per block, plate, and run", 
    format = "latex"
  )
cv$table

```


```{r}

cv<-generate_cv_summary(
    df = filter(yourdata,sample_qc=="PASS"),
    comp_var = parameters$run_var,
    npx = parameters$npx_var,
    threshold =  parameters$Background,
    group_vars = c(parameters$run_var, "plate_id", "block"),
    pivot_var = parameters$run_var,
    value_col = "inter_plate_CV",
    caption = "Average of inter plate CV per block, and run",
    format = "latex"
  )
cv$table

```



```{r}


cv<-generate_cv_summary(
  yourdata, 
  value_col="inter_run_CV",
  comp_var = parameters$run_var, 
  npx = parameters$npx_var, 
  threshold = parameters$Background, 
  group_vars = "block",
  caption="Inter-instrument CV per block", 
  format ="latex") 

cv$table

```


\newpage
\section{CV distribution within block and Instrument}



```{r, echo=FALSE, results='asis'}

  cat("![Intra CV distribution within block and Instrument](", output_dir, "/intracv_regression.", plot_settings$format, "){width=100%}",sep="")

```

\newpage

```{r, echo=FALSE, results='asis'}

  cat("![Inter plate CV distribution within block and Instrument](", output_dir, "/interplatecv_regression.", plot_settings$format, "){width=100%}",sep="")

```


\newpage
\section{Regression of CV across instruments}



```{r, echo=FALSE, results='asis'}

  cat("![Intra CV regression across instruments](", output_dir, "/intra_CV_regression_SC.", plot_settings$format, "){width=100%}",sep="")

```

\newpage

```{r, echo=FALSE, results='asis'}

  cat("![Inter plate CV regression across instruments](", output_dir, "/inter_plate_CV_regression_SC.", plot_settings$format, "){width=100%}",sep="")

```


\newpage
\section{Average detectability per block}


```{r}

results<-summarize_detectability_across_runs(
    df =  filter(yourdata,sample_qc=="PASS"),
    run_var = parameters$run_var,
    npx = parameters$npx_var,
    background = parameters$Background,
    matrix_name = parameters$matrix_column,
    matrix = "Plasma",
    sample_name = "sample_name",
    caption = "Average detectability per block",
    format = "latex"
  )
results$table
```

\newpage
\section{Detectability distribution within block across instruments}


```{r, echo=FALSE, results='asis'}

  cat("![Detectability distribution](", output_dir, "/detectplot.", plot_settings$format, "){width=100%}",sep="")

```



\newpage
\section{Regression of assay detectability across instruments}


```{r, echo=FALSE, results='asis'}

  cat("![Detectability regression across instruments](", output_dir, "/Detectability_plasma_regression_SC.", plot_settings$format, "){width=100%}",sep="")

```



\newpage
\section{Appendix}


\subsection{R Session Information}

```{r, results='asis'}
toLatex(sessionInfo())
```

