---
title: "Product performance report"
author: "Haifa Splittorff"

output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
    includes:
      in_header: ../../templates/olink_style_windows.tex
      before_body: ../../templates/olink_title_style_internal_projects.tex
  html_document:
    df_print: paged
fontsize: 12pt
---
\makeolinktitle
\thispagestyle{olink_header_toc}
\tableofcontents
\newpage
\pagestyle{olink_header}


```{r setup, include=FALSE}
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(fig.pos = 'H',tab.pos = "H", out.extra = "")
#knitr::opts_chunk$set(fig.align="center")
#knitr::opts_chunk$set(fig.height=10,fig.width=12)
knitr::opts_knit$set(eval.after = "fig.cap")

sapply(list.files("../../functions", pattern ="\\.R$", full.names = TRUE), source)

loadfonts(device = "win")

windowsFonts("Arial" = windowsFont("Arial"))
par(family="Arial")
plot(1:10)

options(knitr.table.format = "pandoc")


#Load configuration
config <- yaml::read_yaml("config.yml")

input_data <- config$input_data
parameters <- config$parameters
plot_settings <- config$plot_settings



# Output directory
product_name <- parameters$Product
instrument_name <- parameters$Instrument
project_name <- parameters$Project
background <- parameters$Background
output_dir <- file.path("../../output", product_name, instrument_name)
yourdata <- readRDS(paste0("../../",input_data))%>% mutate(NPX= !!rlang::sym(parameters$npx_var)) # some olink_analyse functions need NPX
matrix_types <- unique((yourdata %>% filter(sample_type=="SAMPLE"))[[as.character(sym(parameters$matrix_column))]] )
outliers<-c()



```

\newpage
\section{Param}

Product :  `r product_name`   
Instrument: `r product_name`   
Project: `r project_name`   
Background: `r background`

\newpage

\section{PCA}


```{r, echo=FALSE, results='asis'}

  cat("![PCA plot for samples using NPX data](", output_dir, "/PCA_plot.", plot_settings$format, "){width=100%}",sep="")

```
\newpage


\section{IQR}


```{r, echo=FALSE, results='asis'}

  cat("![IQR plot for samples using NPX data](", output_dir, "/qc_plot.", plot_settings$format, "){width=100%}",sep="")

```

\newpage


\section{Sample wise NPX distribution}


```{r, echo=FALSE, results='asis'}

  cat("![Sample-wise NPX distribution plate 1](", output_dir, "/distplot_plate1.", plot_settings$format, "){width=100%}",sep="")

```



```{r, echo=FALSE, results='asis'}

  cat("![Sample-wise NPX distribution plate 2](", output_dir, "/distplot_plate2.", plot_settings$format, "){width=100%}",sep="")

```

\newpage


\section{QC summary}

```{r}

generate_qc_table(yourdata, compare_by = parameters$run_var , qc_col = "sample_qc",
                              caption = "QC Status of Samples", font_size = 12, format = "latex")

```



```{r}

generate_qc_table(yourdata, compare_by =parameters$run_var, qc_var = "assay", qc_col = "assay_qc",
                              caption = "QC Status of assays", font_size = 12, format = "latex")

```


\newpage

\section{Total number of counts}

```{r}
generate_count_table_with_block(yourdata, comp_var=parameters$run_var,
                                              values_fill = 0, font_size = 12, format = "latex")
```




\section{Count distribution}



```{r, echo=FALSE, results='asis'}

  cat("![Count distribution](", output_dir, "/count_density.", plot_settings$format, "){width=80%}",sep="")

```




\section{NPX distribution}


```{r, echo=FALSE, results='asis'}

  cat("![NPX distribution](", output_dir, "/NPX_density.", plot_settings$format, "){width=80%}",sep="")

```




<!-- \newpage -->


<!-- \section{Count correlations} -->


<!-- \section{Per assay count correlations} -->

<!-- ```{r, corcount} -->
<!-- result <- correlation_function( -->
<!--   data = filter(yourdata, if (parameters$Background == "detection") { -->
<!--         get(parameters$Background) == "above EOB"   -->
<!--       } else { -->
<!--       NPX > get(parameters$Background) } ,sample_qc=="PASS", !(sample_id%in% outliers) -->
<!--       ), -->
<!--   value_column = "log2count", -->
<!--   comparison_column = parameters$run_var, -->
<!--   filter_sample_type = "SAMPLE", -->
<!--   filter_assay_type = "assay", -->
<!-- ) -->

<!-- create_correlation_table(result$summary , font_size = 8, format = "latex") -->


<!-- ``` -->


<!-- \section{Per assay NPX correlations} -->


<!-- ```{r, NPXcor} -->

<!-- result <- correlation_function( -->
<!--   data = filter(yourdata, if (parameters$Background == "detection") { -->
<!--         get(parameters$Background) == "above EOB"   -->
<!--       } else { -->
<!--       NPX > get(parameters$Background) }), -->
<!--   value_column = "NPX", -->
<!--   comparison_column = parameters$run_var, -->
<!--   filter_sample_type = "SAMPLE", -->
<!--   filter_assay_type = "assay", -->
<!-- ) -->

<!-- create_correlation_table(result$summary , font_size = 8, format = "latex") -->

<!-- ``` -->


\newpage


<!-- \section{Count regression analysis} -->

<!-- ```{r, results='asis', echo=FALSE} -->
<!-- rmd_syntax<- generate_rmd_syntax_for_figures(folder_path=paste0(output_dir,"/regression_plots_log2count_",parameters$run_var), label_prefix = "countreg", width = "100%") -->
<!-- cat(rmd_syntax, sep= "\n\n") -->
<!-- ``` -->




<!-- \newpage -->


<!-- \section{NPX regression analysis} -->

<!-- ```{r, results='asis', echo=FALSE} -->
<!-- rmd_syntax<- generate_rmd_syntax_for_figures(folder_path=paste0(output_dir,"/regression_plots_NPX_",parameters$run_var), label_prefix = "npxreg", width = "100%") -->
<!-- cat(rmd_syntax, sep= "\n\n") -->
<!-- ``` -->

\newpage

\section{External controls}


```{r, echo=FALSE, results='asis'}

  cat("![Sample Control NPX distribution](", output_dir, "/distplot_SAMPLE_CONTROL.", plot_settings$format, "){width=100%}",sep="")

```

\newpage


```{r, echo=FALSE, results='asis'}

  cat("![Negative Control NPX distribution](", output_dir, "/distplot_NEGATIVE_CONTROL.", plot_settings$format, "){width=100%}",sep="")

```

\newpage

```{r, echo=FALSE, results='asis'}

  cat("![Plate Control NPX distribution](", output_dir, "/distplot_PLATE_CONTROL.", plot_settings$format, "){width=100%}",sep="")

```


\newpage
\section{Overall average of intra and inter-plate CV based on sample controls}

```{r}

cv_data<-aggregate_cv(yourdata, "NPX", parameters$Background, comp_var="run_id", filter_value =  "SAMPLE_CONTROL" )

cv_data %>% summarise(intra_CV=mean(intra_CV,na.rm=TRUE), inter_plate_CV=mean(inter_plate_CV, na.rm = TRUE)) %>% 
  mutate(across(everything(), ~ round(.x, digits = 2))) %>% 
   kbl(position = "!h", format = "latex", caption = "Overall average of intra and interCV", longtable = FALSE, booktabs = TRUE)

```


\newpage
\section{Coefficient of variance based on sample controls}


```{r}


cv<-generate_cv_summary(
    df = filter(yourdata,sample_qc=="PASS",!sample_id %in% outliers ),
    comp_var = parameters$run_var,
    npx = parameters$npx_var,
    threshold = parameters$Background,
    group_vars = c(parameters$run_var, "plate_id", "block"),
    pivot_var = parameters$run_var,
    value_col = "intra_CV",
    caption = "Average of intra plate CV per block, plate, and run", format = "latex"
  )
cv$table
```


```{r}

cv<-generate_cv_summary(
    df = filter(yourdata,sample_qc=="PASS",!sample_id %in% outliers),
    comp_var = parameters$run_var,
    npx = parameters$npx_var,
    threshold =  parameters$Background,
    group_vars = c(parameters$run_var, "plate_id", "block"),
    pivot_var = parameters$run_var,
    value_col = "inter_plate_CV",
    caption = "Average of inter plate CV per block, and run", format = "latex"
  )
cv$table
```



```{r}

cv<-generate_cv_summary(
    df = filter(yourdata,sample_qc=="PASS",!sample_id %in% outliers),
    comp_var = parameters$run_var,
    npx = parameters$npx_var,
    threshold =  parameters$Background,
    value_col = "inter_run_CV",
    caption = "Average of inter run CV per block", format = "latex"
  )
cv$table

```


\newpage
\section{Average detectability per block, matrix}

```{r}
process_matrix <- function(matrix_name) {

    results <- summarize_detectability_across_runs(
    df = filter(yourdata, sample_qc == "PASS", !sample_id %in% outliers),
    run_var = parameters$run_var,
    npx = parameters$npx_var,
    background = parameters$Background,
    matrix_name = parameters$matrix_column,
    matrix = matrix_name,
    caption = paste0("Average detectability per block, ", gsub("_", "\\\\_", matrix_name)),
    format = "latex",
    additional_group_vars = "assay_label"
  )

  results$table

}

 
```



```{r, results='asis', echo=FALSE}
for (mat in matrix_types) {
  cat(process_matrix(mat), "\n\n")
}

 
```



\newpage
\section{ Concordance Detectability distribution within block across replicate runs}


```{r, echo=FALSE, results='asis'}

  cat("![Detectability distribution](", output_dir, "/detectplot_concordance.", plot_settings$format, "){width=100%}",sep="")

```


\newpage

\section{Percentage of assays detectable in at least 10, 50 percent of the concordance samples}


```{r}

results<-summarize_detectability_across_runs(
  df = filter(yourdata,sample_qc=="PASS", !sample_id %in% outliers, sample_label=="concordance"),
  run_var = parameters$run_var,
  npx = parameters$npx_var,
  background = parameters$Background,
  caption = "Average detectability per block",
  format = "latex",
  additional_group_vars = "assay_label"
)

calculate_detectability_percentage (results$data, detectability_col= "median_detectability" , group_var = "assay_label", thresholds = c(0.1,0.5)) %>%
  kbl(position = "!h", format="latex", caption = "Percentage of assays detectable in at least 10, 50 percent of the concordance samples", ) %>%
  kable_classic(full_width = FALSE)

```

\newpage
\section{Percentage of assays detectable in at least one sample}


```{r, results='asis', echo=FALSE}

process_matrix2 <- function(matrix_name) {

    results <- summarize_detectability_across_runs(
    df = filter(yourdata, sample_qc == "PASS", !sample_id %in% outliers),
    run_var = parameters$run_var,
    npx = parameters$npx_var,
    background = parameters$Background,
    matrix_name = parameters$matrix_column,
    matrix = matrix_name,
    caption = paste0("Average detectability per block, ", matrix_name),
    format = "latex",
    additional_group_vars = "assay_label"
  )

  calculate_detectability_percentage(
    results$data,
    detectability_col = "median_detectability",
    group_var = "assay_label",
    thresholds = c(0)
  ) %>%
    kbl(position = "!h", format = "latex",
        caption = paste0("Percentage of assays detectable in at least one sample in ", gsub("_", "\\\\_", matrix_name), " samples"),
        col.names = c("Group","N", "Percentage")) %>%
    kable_classic(full_width = FALSE)

}

for (mat in matrix_types) {
  cat(process_matrix2(mat), "\n\n")
}

```




\newpage
\section{Overlap of detectable assays across matrices}


```{r, echo=FALSE, results='asis'}

  cat("![Overlap of assays detectable in at least one sample in all 3 replicate runs, across matrices](", output_dir, "/Venn_detect_across_matrix.", plot_settings$format, "){width=100%}",sep="")

```


\newpage
\section{Appendix}


\subsection{R Session Information}

```{r, results='asis'}
toLatex(sessionInfo())
```

