---
title: "Lysate module report"
author: "Haifa Splittorff"

output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
    includes:
      in_header: templates/olink_style_windows.tex
      before_body: templates/olink_title_style_internal_projects.tex
  html_document:
    df_print: paged
fontsize: 12pt
---
\makeolinktitle
\thispagestyle{olink_header_toc}
\tableofcontents
\newpage
\pagestyle{olink_header}


```{r setup, include=FALSE}
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(fig.pos = 'H',tab.pos = "H", out.extra = "")
knitr::opts_knit$set(eval.after = "fig.cap")

source("Rcode/mylib.R")
sapply(list.files("Functions", pattern ="\\.R$", full.names = TRUE), source)

loadfonts(device = "win")

windowsFonts("Arial" = windowsFont("Arial"))
par(family="Arial")
plot(1:10)

options(knitr.table.format = "pandoc")


# Load configuration
config <- yaml::read_yaml("config_product.yml")

input_data <- config$input_lysate
parameters <- config$parameters
plot_settings <- config$plot_settings


# Output directory
product_name <- parameters$Product
project_name <- parameters$Project
output_dir <- file.path("Results", product_name, "Lysate_module")
outliers<-c()
yourdata <- read_rds(input_data) %>% mutate(NPX= !!rlang::sym(parameters$npx_var)) %>% # some olink_analyse functions need NPX
filter(sample_qc=="PASS", !(sample_id %in% outliers))                                         
matrix_types <- unique(yourdata[[as.character(sym(parameters$matrix_column))]] )



```


\newpage
\section{Param}

Product :  `r product_name`   
Project: `r project_name`   

<!-- \newpage -->

<!-- \section{PCA} -->


<!-- ```{r, echo=FALSE, results='asis'} -->

<!--   cat("![PCA plot for samples using NPX data](", output_dir, "/PCA_plot.", plot_settings$format, "){width=100%}",sep="") -->

<!-- ``` -->
<!-- \newpage -->


<!-- \section{IQR} -->


<!-- ```{r, echo=FALSE, results='asis'} -->

<!--   cat("![IQR plot for samples using NPX data](", output_dir, "/qc_plot.", plot_settings$format, "){width=100%}",sep="") -->

<!-- ``` -->


<!-- \newpage -->


<!-- \section{Sample wise NPX distribution} -->


<!-- ```{r, echo=FALSE, results='asis'} -->

<!--   cat("![Sample-wise NPX distribution](", output_dir, "/distplot.", plot_settings$format, "){width=100%}",sep="") -->

<!-- ``` -->





<!-- \newpage -->
<!-- \section{Count density, based only on concordance samples} -->



<!-- ```{r, echo=FALSE, results='asis'} -->

<!--   cat("![Count distribution](", output_dir, "/count_density.", plot_settings$format, "){width=80%}",sep="") -->

<!-- ``` -->




<!-- \section{NPX density, based only on concordance samples} -->


<!-- ```{r, echo=FALSE, results='asis'} -->

<!--   cat("![NPX distribution](", output_dir, "/NPX_density.", plot_settings$format, "){width=80%}",sep="") -->

<!-- ``` -->





<!-- \newpage -->


<!-- \section{Count regression analysis} -->

<!-- ```{r, results='asis', echo=FALSE} -->
<!--   cat("![Log2count regression](", output_dir, "/log2count_regression.", plot_settings$format, "){width=100%}",sep="") -->

<!-- ``` -->




<!-- \newpage -->


<!-- \section{NPX regression analysis} -->

<!-- ```{r, results='asis', echo=FALSE} -->
<!--   cat("![NPX regression](", output_dir, "/NPX_regression.", plot_settings$format, "){width=100%}",sep="") -->

<!-- ``` -->

<!-- \newpage -->

<!-- \section{Per assay count correlations} -->
<!-- \subsection{Including all data points} -->


<!-- ```{r, corcount} -->
<!-- count_cor <- correlation_function( -->
<!--     data = Concordance_count_npx %>% mutate(plate_id="all_plates",block ="all_block", sample_index=sample_name), -->
<!--     value_column = "log2count", -->
<!--     comparison_column = parameters$run_var, -->
<!--     filter_sample_type = "SAMPLE", -->
<!--     filter_assay_type = "assay" -->
<!-- ) -->

<!-- create_correlation_table(count_cor$summary , font_size = 12, format = "latex") -->


<!-- ``` -->


<!-- \subsection{Including only data points above product specific background} -->

<!-- ```{r, corcount2} -->
<!-- count_cor <- correlation_function( -->
<!--     data = Concordance_count_npx_above_background %>% mutate(plate_id="all_plates",block ="all_block", sample_index=sample_name), -->
<!--     value_column = "log2count", -->
<!--     comparison_column = parameters$run_var, -->
<!--     filter_sample_type = "SAMPLE", -->
<!--     filter_assay_type = "assay" -->
<!-- ) -->

<!-- create_correlation_table(count_cor$summary , font_size = 12, format = "latex") -->


<!-- ``` -->

<!-- \newpage -->

<!-- \section{Per assay NPX correlations} -->

<!-- \subsection{Including all data points} -->


<!-- ```{r, NPXcor} -->

<!-- NPX_cor <- correlation_function( -->
<!--     data = Concordance_count_npx %>% mutate(plate_id="all_plates",block ="all_block", sample_index=sample_name), -->
<!--     value_column = "NPX", -->
<!--     comparison_column = parameters$run_var, -->
<!--     filter_sample_type = "SAMPLE", -->
<!--     filter_assay_type = "assay" -->
<!-- ) -->

<!-- create_correlation_table(NPX_cor$summary , font_size = 12, format = "latex") -->

<!-- ``` -->


<!-- \subsection{Including only data points above background} -->


<!-- ```{r, NPXcor2} -->

<!-- NPX_cor <- correlation_function( -->
<!--     data = Concordance_count_npx_above_background %>% mutate(plate_id="all_plates",block ="all_block", sample_index=sample_name), -->
<!--     value_column = "NPX", -->
<!--     comparison_column = parameters$run_var, -->
<!--     filter_sample_type = "SAMPLE", -->
<!--     filter_assay_type = "assay" -->
<!-- ) -->

<!-- create_correlation_table(NPX_cor$summary , font_size = 12, format = "latex") -->

<!-- ``` -->



<!-- \newpage -->

<!-- \section{Per assay NPX correlations distribution} -->



<!-- ```{r, results='asis', echo=FALSE} -->
<!--   cat("![NPX correlation distribution](", output_dir, "/corrplot.", plot_settings$format, "){width=100%}",sep="") -->

<!-- ``` -->




\newpage
\section{Coefficient of variance based on Lysate replicate samples}

Replicates are different on different plates, interplate CV can not be calculated.  
For each product, cv values from all replicate runs were aggregated to one single value using the median.  

\subsection{CV based on tissue lysate replicates}


```{r}
cv_data<-aggregate_cv(yourdata, "NPX", parameters$Background, comp_var="run_id", filter_col = "sample_label", filter_value =  "Tissue_Lysate_replicates", additionnal_var=c(parameters$run_var,"block"))

cv_data %>% group_by(!!sym(parameters$run_var)) %>% summarise(Median_intraCV=median(intra_CV,na.rm=TRUE),
                                                 Median_inter_runCV=median(inter_run_CV,na.rm=TRUE), .groups = "drop") %>%  mutate(across(where(is.numeric), ~ round(., digits = 2))) %>% 
kbl(position = "!h",format="latex", booktabs = T, caption = "Overall Median of CV based on Tissue Lysate replicates") %>% kable_styling(font_size = 12,full_width = FALSE)

```



```{r}

cv_data %>% group_by(!!sym(parameters$run_var),block) %>% summarise(Mean_intraCV=mean(intra_CV, na.rm = TRUE),
                                                                             Median_intraCV=median(intra_CV,na.rm=TRUE),
                                                                             Mean_inter_runCV=mean(inter_run_CV, na.rm = TRUE),
                                                                             Median_inter_runCV=median(inter_run_CV,na.rm=TRUE), .groups = "drop") %>%  mutate(across(where(is.numeric), ~ round(., digits = 2))) %>% 
kbl(position = "!h",format="latex", booktabs = T, caption = "Mean and Median of CV based on Tissue Lysate replicates, per block") %>% kable_styling(font_size = 12,full_width = FALSE)

```


\newpage


```{r, results='asis', echo=FALSE}
  cat("![Distribution of intraCV based on tissue lysate replicates](", output_dir, "/violin_intra_tissue.", plot_settings$format, "){width=100%}",sep="")

```

\newpage
\subsection{CV based on cell lysate replicates}


```{r}
cv_data<-aggregate_cv(yourdata, "NPX", parameters$Background, comp_var="run_id", filter_col = "sample_label", filter_value =  "Cell_Lysate_replicates", additionnal_var=c(parameters$run_var,"block"))

cv_data %>% group_by(!!sym(parameters$run_var)) %>% summarise(Median_intraCV=median(intra_CV,na.rm=TRUE),
                                                 Median_inter_runCV=median(inter_run_CV,na.rm=TRUE), .groups = "drop") %>%  mutate(across(where(is.numeric), ~ round(., digits = 2))) %>% 
kbl(position = "!h",format="latex", booktabs = T, caption = "Overall Median of CV based on Cell Lysate replicates") %>% kable_styling(font_size = 12,full_width = FALSE)

```



```{r}


cv_data %>% group_by(!!sym(parameters$run_var),block) %>% summarise(Mean_intraCV=mean(intra_CV, na.rm = TRUE),
                                                                             Median_intraCV=median(intra_CV,na.rm=TRUE),
                                                                             Mean_inter_runCV=mean(inter_run_CV, na.rm = TRUE),
                                                                             Median_inter_runCV=median(inter_run_CV,na.rm=TRUE), .groups = "drop") %>%  mutate(across(where(is.numeric), ~ round(., digits = 2))) %>% 
kbl(position = "!h",format="latex", booktabs = T, caption = "Mean and Median of CV based on Cell Lysate replicates,  per block") %>% kable_styling(font_size = 12,full_width = FALSE)

```



\newpage


```{r, results='asis', echo=FALSE}
  cat("![Distribution of intraCV based on tissue lysate replicates](", output_dir, "/violin_intra_cell.", plot_settings$format, "){width=100%}",sep="")

```




\newpage
\section{Average detectability per replicate run}



```{r}

detectability <- summarize_detectability_across_runs(
    df = filter(yourdata),
    run_var =  "run_id",
    npx = parameters$npx_var,
    background = parameters$Background,
    caption = "Average detectability per replicate run",
    format = "latex",
    additional_group_vars = parameters$run_var
  )

detectability$data %>% group_by(!!sym(parameters$run_var)) %>% summarise(Rep1=round(mean(rep1, na.rm = TRUE),digits = 3),
                                                                    Rep2=round(mean(rep2, na.rm = TRUE),digits = 3),
                                                                    Rep3=round(mean(rep3, na.rm = TRUE),digits = 3)
                                                                    ) %>% 
kbl(position = "!h",format="latex", booktabs = T, caption = "Average detectability per replicate run and product") %>% kable_styling(font_size = 12,full_width = FALSE)

```




\newpage
\section{Percentage of assays detectable in at least one lysate sample}


```{r, echo=FALSE, results='asis'}

detectability_data<-aggregate_detectability(
df = yourdata,
sample_name = "sample_name",
comp_var = parameters$run_var,
run_var = "run_id",
npx = parameters$npx_var,
background = parameters$Background
)

calculate_detectability_percentage_per_group(
     df = detectability_data, 
     assay_group = unique(yourdata %>% select(olink_id,assay_label)), 
     detectability_col = "Detectability", 
 thresholds = c( 0),
   group_var = "assay_label",
   instrument_col = "Product") %>% kbl(position = "!h", format="latex", caption = "Percentage of assays detectable in at least one sample per product" ) %>%
        kable_classic(full_width = FALSE)

```



\newpage
\section{Percentage of assays detectable in at least 10, 50 percent of the lysate samples}


```{r, echo=FALSE, results='asis'}

calculate_detectability_percentage_per_group(
     df = detectability_data, 
     assay_group = unique(yourdata %>% select(olink_id,assay_label)), 
     detectability_col = "Detectability", 
 thresholds = c( 0.1, 0.5),
   group_var = "assay_label",
   instrument_col = "Product") %>% kbl(position = "!h", format="latex", caption = "Percentage of assays detectable in at least 10, 50 percent of the lysate samples" ) %>%
        kable_classic(full_width = FALSE)

```




\newpage
\section{Overlap of assays detectable in at least 10, 50 percent of the lysate samples between products}


```{r, results='asis', echo=FALSE}
  cat("![Overlap of assays detectable in at least 10 and 50 percent between products](", output_dir, "/Venn_detect.", plot_settings$format, "){width=100%}",sep="")

```


}




\newpage
\section{Appendix}


\subsection{R Session Information}

```{r, results='asis'}
toLatex(sessionInfo())
```

