---
title: "Lysate module report"
author: "Haifa Splittorff"

output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
    includes:
      in_header: ../../templates/olink_style_windows.tex
      before_body: ../../templates/olink_title_style_internal_projects.tex
  html_document:
    df_print: paged
fontsize: 12pt
---
\makeolinktitle
\thispagestyle{olink_header_toc}
\tableofcontents
\newpage
\pagestyle{olink_header}


```{r setup, include=FALSE}
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(fig.pos = 'H',tab.pos = "H", out.extra = "")
knitr::opts_knit$set(eval.after = "fig.cap")

sapply(list.files("../../functions", pattern ="\\.R$", full.names = TRUE), source)

loadfonts(device = "win")

windowsFonts("Arial" = windowsFont("Arial"))
par(family="Arial")
plot(1:10)

options(knitr.table.format = "pandoc")


# Load configuration
#config <- yaml::read_yaml("config.yml")

input_data <- config$input_data
parameters <- config$parameters
plot_settings <- config$plot_settings


# Output directory
product_name <- parameters$Product
project_name <- parameters$Project
background <- parameters$Background
output_dir <- file.path("../../output", product_name, "Lysate_module")
outliers<-c()
yourdata <- read_rds(paste0("../../",input_data)) %>% mutate(NPX= !!rlang::sym(parameters$npx_var)) %>% # some olink_analyse functions need NPX
filter(sample_qc=="PASS", !(sample_id %in% outliers))                                         
matrix_types <- unique(yourdata[[as.character(sym(parameters$matrix_column))]] )



```


\newpage
\section{Param}

Product :  `r product_name`   
Project: `r project_name`   
Background: `r background`


\newpage
\section{Coefficient of variance based on Lysate replicate samples}

Replicates are different on different plates, interplate CV can not be calculated.  
For each product, cv values from all replicate runs were aggregated to one single value using the median.  

\subsection{CV based on tissue lysate replicates pool 1}


```{r}
cv_data<-aggregate_cv(yourdata, "NPX", parameters$Background, comp_var="run_id", filter_col = "sample_label", filter_value =  "Tissue_Lysate_replicates_pool1", additionnal_var=c(parameters$run_var,"block"))

cv_data %>% group_by(!!sym(parameters$run_var)) %>% summarise(Median_intraCV=median(intra_CV,na.rm=TRUE),
                                                 Median_inter_runCV=median(inter_run_CV,na.rm=TRUE), .groups = "drop") %>%  mutate(across(where(is.numeric), ~ round(., digits = 2))) %>% 
kbl(position = "!h",format="latex", booktabs = T, caption = "Overall Median of CV based on Tissue Lysate replicates pool 1") %>% kable_styling(font_size = 12,full_width = FALSE)

```



```{r}

cv_data %>% group_by(!!sym(parameters$run_var),block) %>% summarise(Mean_intraCV=mean(intra_CV, na.rm = TRUE),
                                                                             Median_intraCV=median(intra_CV,na.rm=TRUE),
                                                                             Mean_inter_runCV=mean(inter_run_CV, na.rm = TRUE),
                                                                             Median_inter_runCV=median(inter_run_CV,na.rm=TRUE), .groups = "drop") %>%  mutate(across(where(is.numeric), ~ round(., digits = 2))) %>% 
kbl(position = "!h",format="latex", booktabs = T, caption = "Mean and Median of CV based on Tissue Lysate replicates pool 1, per block") %>% kable_styling(font_size = 12,full_width = FALSE)

```


\newpage


```{r, results='asis', echo=FALSE}
  cat("![Distribution of intraCV based on tissue lysate replicates pool 1](", output_dir, "/violin_intra_tissue_pool1.", plot_settings$format, "){width=100%}",sep="")

```


\subsection{CV based on tissue lysate replicates pool 2}


```{r}
cv_data<-aggregate_cv(yourdata, "NPX", parameters$Background, comp_var="run_id", filter_col = "sample_label", filter_value =  "Tissue_Lysate_replicates_pool2", additionnal_var=c(parameters$run_var,"block"))

cv_data %>% group_by(!!sym(parameters$run_var)) %>% summarise(Median_intraCV=median(intra_CV,na.rm=TRUE),
                                                 Median_inter_runCV=median(inter_run_CV,na.rm=TRUE), .groups = "drop") %>%  mutate(across(where(is.numeric), ~ round(., digits = 2))) %>% 
kbl(position = "!h",format="latex", booktabs = T, caption = "Overall Median of CV based on Tissue Lysate replicates pool 2") %>% kable_styling(font_size = 12,full_width = FALSE)

```



```{r}

cv_data %>% group_by(!!sym(parameters$run_var),block) %>% summarise(Mean_intraCV=mean(intra_CV, na.rm = TRUE),
                                                                             Median_intraCV=median(intra_CV,na.rm=TRUE),
                                                                             Mean_inter_runCV=mean(inter_run_CV, na.rm = TRUE),
                                                                             Median_inter_runCV=median(inter_run_CV,na.rm=TRUE), .groups = "drop") %>%  mutate(across(where(is.numeric), ~ round(., digits = 2))) %>% 
kbl(position = "!h",format="latex", booktabs = T, caption = "Mean and Median of CV based on Tissue Lysate replicates pool 2, per block") %>% kable_styling(font_size = 12,full_width = FALSE)

```


\newpage


```{r, results='asis', echo=FALSE}
  cat("![Distribution of intraCV based on tissue lysate replicates pool 2](", output_dir, "/violin_intra_tissue_pool2.", plot_settings$format, "){width=100%}",sep="")

```


\newpage
\subsection{CV based on cell lysate replicates pool 1}


```{r}
cv_data<-aggregate_cv(yourdata, "NPX", parameters$Background, comp_var="run_id", filter_col = "sample_label", filter_value =  "Cell_Lysate_replicates_pool1", additionnal_var=c(parameters$run_var,"block"))

cv_data %>% group_by(!!sym(parameters$run_var)) %>% summarise(Median_intraCV=median(intra_CV,na.rm=TRUE),
                                                 Median_inter_runCV=median(inter_run_CV,na.rm=TRUE), .groups = "drop") %>%  mutate(across(where(is.numeric), ~ round(., digits = 2))) %>% 
kbl(position = "!h",format="latex", booktabs = T, caption = "Overall Median of CV based on Cell Lysate replicates pool 1") %>% kable_styling(font_size = 12,full_width = FALSE)

```



```{r}


cv_data %>% group_by(!!sym(parameters$run_var),block) %>% summarise(Mean_intraCV=mean(intra_CV, na.rm = TRUE),
                                                                             Median_intraCV=median(intra_CV,na.rm=TRUE),
                                                                             Mean_inter_runCV=mean(inter_run_CV, na.rm = TRUE),
                                                                             Median_inter_runCV=median(inter_run_CV,na.rm=TRUE), .groups = "drop") %>%  mutate(across(where(is.numeric), ~ round(., digits = 2))) %>% 
kbl(position = "!h",format="latex", booktabs = T, caption = "Mean and Median of CV based on Cell Lysate replicates pool 1,  per block") %>% kable_styling(font_size = 12,full_width = FALSE)

```



\newpage


```{r, results='asis', echo=FALSE}
  cat("![Distribution of intraCV based on tissue lysate replicates pool 1](", output_dir, "/violin_intra_cell_pool1.", plot_settings$format, "){width=100%}",sep="")

```



\newpage
\subsection{CV based on cell lysate replicates pool 2}


```{r}
cv_data<-aggregate_cv(yourdata, "NPX", parameters$Background, comp_var="run_id", filter_col = "sample_label", filter_value =  "Cell_Lysate_replicates_pool2", additionnal_var=c(parameters$run_var,"block"))

cv_data %>% group_by(!!sym(parameters$run_var)) %>% summarise(Median_intraCV=median(intra_CV,na.rm=TRUE),
                                                 Median_inter_runCV=median(inter_run_CV,na.rm=TRUE), .groups = "drop") %>%  mutate(across(where(is.numeric), ~ round(., digits = 2))) %>% 
kbl(position = "!h",format="latex", booktabs = T, caption = "Overall Median of CV based on Cell Lysate replicates pool 2") %>% kable_styling(font_size = 12,full_width = FALSE)

```



```{r}


cv_data %>% group_by(!!sym(parameters$run_var),block) %>% summarise(Mean_intraCV=mean(intra_CV, na.rm = TRUE),
                                                                             Median_intraCV=median(intra_CV,na.rm=TRUE),
                                                                             Mean_inter_runCV=mean(inter_run_CV, na.rm = TRUE),
                                                                             Median_inter_runCV=median(inter_run_CV,na.rm=TRUE), .groups = "drop") %>%  mutate(across(where(is.numeric), ~ round(., digits = 2))) %>% 
kbl(position = "!h",format="latex", booktabs = T, caption = "Mean and Median of CV based on Cell Lysate replicates pool 2,  per block") %>% kable_styling(font_size = 12,full_width = FALSE)

```



\newpage


```{r, results='asis', echo=FALSE}
  cat("![Distribution of intraCV based on tissue lysate replicates pool 2](", output_dir, "/violin_intra_cell_pool2.", plot_settings$format, "){width=100%}",sep="")

```



\newpage
\section{Average detectability per replicate run}



```{r}

detectability <- summarize_detectability_across_runs(
    df = filter(yourdata),
    run_var =  "run_id",
    npx = parameters$npx_var,
    background = parameters$Background,
    caption = "Average detectability per replicate run",
    format = "latex",
    additional_group_vars = parameters$run_var
  )

detectability$data %>% group_by(!!sym(parameters$run_var)) %>% summarise(Rep1=round(mean(rep1, na.rm = TRUE),digits = 3),
                                                                    Rep2=round(mean(rep2, na.rm = TRUE),digits = 3),
                                                                    Rep3=round(mean(rep3, na.rm = TRUE),digits = 3)
                                                                    ) %>% 
kbl(position = "!h",format="latex", booktabs = T, caption = "Average detectability per replicate run and product") %>% kable_styling(font_size = 12,full_width = FALSE)

```




\newpage
\section{Percentage of assays detectable in at least one lysate sample}


```{r, echo=FALSE, results='asis'}

detectability_data<-aggregate_detectability(
df = yourdata,
sample_name = "sample_name",
comp_var = parameters$run_var,
run_var = "run_id",
npx = parameters$npx_var,
background = parameters$Background
)

calculate_detectability_percentage_per_group(
     df = detectability_data, 
     assay_group = unique(yourdata %>% select(olink_id,assay_label)), 
     detectability_col = "Detectability", 
 thresholds = c( 0),
   group_var = "assay_label",
   instrument_col = "Product") %>% kbl(position = "!h", format="latex", caption = "Percentage of assays detectable in at least one sample per product" ) %>%
        kable_classic(full_width = FALSE)

```



\newpage
\section{Percentage of assays detectable in at least 10, 50 percent of the lysate samples}


```{r, echo=FALSE, results='asis'}

calculate_detectability_percentage_per_group(
     df = detectability_data, 
     assay_group = unique(yourdata %>% select(olink_id,assay_label)), 
     detectability_col = "Detectability", 
 thresholds = c( 0.1, 0.5),
   group_var = "assay_label",
   instrument_col = "Product") %>% kbl(position = "!h", format="latex", caption = "Percentage of assays detectable in at least 10, 50 percent of the lysate samples" ) %>%
        kable_classic(full_width = FALSE)

```




\newpage
\section{Overlap of assays detectable in at least 10, 50 percent of the lysate samples between products}


```{r, results='asis', echo=FALSE}
  cat("![Overlap of assays detectable in at least 10 and 50 percent between products](", output_dir, "/Venn_detect.", plot_settings$format, "){width=100%}",sep="")

```


}




\newpage
\section{Appendix}


\subsection{R Session Information}

```{r, results='asis'}
toLatex(sessionInfo())
```

