---
title: "Product comparison report"
author: "Haifa Splittorff"

output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
    includes:
      in_header: ../../templates/olink_style_windows.tex
      before_body: ../../templates/olink_title_style_internal_projects.tex
  html_document:
    df_print: paged
fontsize: 12pt
---
\makeolinktitle
\thispagestyle{olink_header_toc}
\tableofcontents
\newpage
\pagestyle{olink_header}



```{r setup, include=FALSE}
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(fig.pos = 'H',tab.pos = "H", out.extra = "")
knitr::opts_knit$set(eval.after = "fig.cap")

sapply(list.files("../../functions", pattern ="\\.R$", full.names = TRUE), source)

loadfonts(device = "win")

windowsFonts("Arial" = windowsFont("Arial"))
par(family="Arial")
plot(1:10)

options(knitr.table.format = "pandoc")


# Load configuration
config <- yaml::read_yaml("config.yml")

input_data <- config$input_data
parameters <- config$parameters
plot_settings <- config$plot_settings


# Output directory
product_name <- parameters$Product
project_name <- parameters$Project
background <- parameters$Background
output_dir <- file.path("../../output", product_name, "Product_comparison")
outliers<-c()
yourdata <- read_rds(paste0("../../",input_data)) %>% mutate(NPX= !!rlang::sym(parameters$npx_var)) %>% # some olink_analyse functions need NPX
filter(sample_qc=="PASS", !(sample_id %in% outliers))                                  
Concordance_count_npx_above_background<-read_rds("Concordance_count_npx_above_background.rds")
Concordance_count_npx<-read_rds("Concordance_count_npx.rds")
matrix_types <- unique((yourdata %>% filter(sample_type=="SAMPLE"))[[as.character(sym(parameters$matrix_column))]] )


```


\newpage
\section{Param}

Product :  `r product_name`   
Project: `r project_name`   
Background: `r background`

<!-- \newpage -->

<!-- \section{PCA} -->


<!-- ```{r, echo=FALSE, results='asis'} -->

<!--   cat("![PCA plot for samples using NPX data](", output_dir, "/PCA_plot.", plot_settings$format, "){width=100%}",sep="") -->

<!-- ``` -->
<!-- \newpage -->


<!-- \section{IQR} -->


<!-- ```{r, echo=FALSE, results='asis'} -->

<!--   cat("![IQR plot for samples using NPX data](", output_dir, "/qc_plot.", plot_settings$format, "){width=100%}",sep="") -->

<!-- ``` -->



\newpage
\section{Summary of the PRS metrics}


\subsection{Median of the correlations between products}

```{r}

NPX_cor <- correlation_function(
    data = Concordance_count_npx_above_background %>% mutate(plate_id="all_plates",block ="all_block", sample_index=sample_name),
    value_column = "NPX",
    comparison_column = parameters$run_var,
    filter_sample_type = "SAMPLE",
    filter_assay_type = "assay"
)

```


|Metric | formula|Median of the differences|
|-------------|---------------------|----------------|
|NPX correlations| per assay correlation| `r round(NPX_cor$summary[["Median"]],digits=3)` |



\subsection{Median of the differences in CV}

```{r}
cv_data<-aggregate_cv(yourdata, "NPX", parameters$Background, comp_var="run_id", filter_value =  "SAMPLE_CONTROL", additionnal_var=parameters$run_var) 

inter<-compute_median_diff(cv_data,"Product","inter_plate_CV")
intra<- compute_median_diff(cv_data,"Product","intra_CV")

```


|Metric |Difference formula|Median of the differences|
|-------------|---------------------|----------------|
|IntraCV| `r intra[[1]]`| `r round(intra[[2]],digits=3)` |
|InterCV| `r inter[[1]]`| `r round(inter[[2]],digits=3)` |


\subsection{Median of the differences in detectability}

```{r}
detectability_data<-aggregate_detectability(
    df = filter(yourdata,sample_qc=="PASS", !(sample_id%in% outliers), sample_label=="concordance"),
    sample_name = "sample_name",
    comp_var = parameters$run_var,
    run_var = "run_id",
    npx = parameters$npx_var,
    background = parameters$Background,
    )

detect<-compute_median_diff(detectability_data,"Product","Detectability")
```


|Metric |Difference formula|Median of the differences|
|-------------|---------------------|----------------|
|Detectability| `r detect[[1]]`| `r round(detect[[2]],digits=3)` |



\newpage


\section{Sample wise NPX distribution}


```{r, echo=FALSE, results='asis'}

  cat("![Sample-wise NPX distribution](", output_dir, "/distplot.", plot_settings$format, "){width=100%}",sep="")

```



\newpage
\section{Count distribution within categories, based only on concordance samples}

Only overlapping assays are included below


```{r}

result<- descriptive_tabulation(filter(yourdata %>% filter(!is.na(ht_oid)), sample_label=="concordance"),parameters$run_var)

result$count_categ_table %>% kbl(position = "!h",  caption="Count distribution within categories")

```



\newpage
\section{Count density, based only on concordance samples}



```{r, echo=FALSE, results='asis'}

  cat("![Count distribution](", output_dir, "/count_density.", plot_settings$format, "){width=80%}",sep="")

```




\section{NPX density, based only on concordance samples}


```{r, echo=FALSE, results='asis'}

  cat("![NPX distribution](", output_dir, "/NPX_density.", plot_settings$format, "){width=80%}",sep="")

```





\newpage


\section{Count regression analysis}

```{r, results='asis', echo=FALSE}
  cat("![Log2count regression](", output_dir, "/log2count_regression.", plot_settings$format, "){width=100%}",sep="")

```




\newpage


\section{NPX regression analysis}

```{r, results='asis', echo=FALSE}
  cat("![NPX regression](", output_dir, "/NPX_regression.", plot_settings$format, "){width=100%}",sep="")

```

\newpage

\section{Per assay count correlations}
\subsection{Including all data points}


```{r, corcount}
count_cor <- correlation_function(
    data = Concordance_count_npx %>% mutate(plate_id="all_plates",block ="all_block", sample_index=sample_name),
    value_column = "log2count",
    comparison_column = parameters$run_var,
    filter_sample_type = "SAMPLE",
    filter_assay_type = "assay"
)

create_correlation_table(count_cor$summary , font_size = 12, format = "latex")


```



```{r, corcountb}
count_cor <- correlation_function(
    data = Concordance_count_npx %>% mutate(plate_id="all_plates", sample_index=sample_name),
    value_column = "log2count",
    comparison_column = parameters$run_var,
    filter_sample_type = "SAMPLE",
    filter_assay_type = "assay"
)

create_correlation_table(count_cor$summary , font_size = 12, format = "latex")


```


\newpage
\subsection{Including only data points above product specific background}

```{r, corcount2}
count_cor <- correlation_function(
    data = Concordance_count_npx_above_background %>% mutate(plate_id="all_plates",block ="all_block", sample_index=sample_name),
    value_column = "log2count",
    comparison_column = parameters$run_var,
    filter_sample_type = "SAMPLE",
    filter_assay_type = "assay"
)

create_correlation_table(count_cor$summary , font_size = 12, format = "latex")


```


```{r, corcount2b}
count_cor <- correlation_function(
    data = Concordance_count_npx_above_background %>% mutate(plate_id="all_plates", sample_index=sample_name),
    value_column = "log2count",
    comparison_column = parameters$run_var,
    filter_sample_type = "SAMPLE",
    filter_assay_type = "assay"
)

create_correlation_table(count_cor$summary , font_size = 12, format = "latex")


```

\newpage

\section{Per assay NPX correlations}

\subsection{Including all data points}


```{r, NPXcor}

NPX_cor <- correlation_function(
    data = Concordance_count_npx %>% mutate(plate_id="all_plates",block ="all_block", sample_index=sample_name),
    value_column = "NPX",
    comparison_column = parameters$run_var,
    filter_sample_type = "SAMPLE",
    filter_assay_type = "assay"
)

create_correlation_table(NPX_cor$summary , font_size = 12, format = "latex")

```


```{r, NPXcorb}

NPX_cor <- correlation_function(
    data = Concordance_count_npx %>% mutate(plate_id="all_plates", sample_index=sample_name),
    value_column = "NPX",
    comparison_column = parameters$run_var,
    filter_sample_type = "SAMPLE",
    filter_assay_type = "assay"
)

create_correlation_table(NPX_cor$summary , font_size = 12, format = "latex")

```

\subsection{Including only data points above background}


```{r, NPXcor2}

NPX_cor <- correlation_function(
    data = Concordance_count_npx_above_background %>% mutate(plate_id="all_plates",block ="all_block", sample_index=sample_name),
    value_column = "NPX",
    comparison_column = parameters$run_var,
    filter_sample_type = "SAMPLE",
    filter_assay_type = "assay"
)

create_correlation_table(NPX_cor$summary , font_size = 12, format = "latex")

```


```{r, NPXcor2b}

NPX_cor <- correlation_function(
    data = Concordance_count_npx_above_background %>% mutate(plate_id="all_plates", sample_index=sample_name),
    value_column = "NPX",
    comparison_column = parameters$run_var,
    filter_sample_type = "SAMPLE",
    filter_assay_type = "assay"
)

create_correlation_table(NPX_cor$summary , font_size = 12, format = "latex")

```



\newpage

\section{Per assay NPX correlations distribution}

\subsection{All data points included}


```{r, results='asis', echo=FALSE}
  cat("![NPX correlation distribution](", output_dir, "/corrplot_all.", plot_settings$format, "){width=100%}",sep="")

```


\subsection{Only data above product specific background included}


```{r, results='asis', echo=FALSE}
  cat("![NPX correlation distribution](", output_dir, "/corrplot_above.", plot_settings$format, "){width=100%}",sep="")

```


\newpage
\section{Coefficient of variance based on samples controls}


```{r}

cv_data<-aggregate_cv(yourdata, "NPX", parameters$Background, comp_var="run_id", filter_value =  "SAMPLE_CONTROL", additionnal_var=c(parameters$run_var, "assay_label"))

  cv_data %>% group_by(!!sym(parameters$run_var), assay_label) %>% summarise(Mean_intraCV=mean(intra_CV, na.rm = TRUE),
                                                                             Median_intraCV=median(intra_CV,na.rm=TRUE),
                                                                             Mean_interCV=mean(inter_plate_CV, na.rm = TRUE),
                                                                             Median_interCV=median(inter_plate_CV,na.rm=TRUE), .groups = "drop") %>%  mutate(across(where(is.numeric), ~ round(., digits = 2))) %>% 
kbl(position = "!h",format="latex", booktabs = T, caption = "Mean and Median of CV based on samples controls") %>% kable_styling(font_size = 12,full_width = FALSE)

```





```{r}

cv_data<-aggregate_cv(yourdata, "NPX", parameters$Background, comp_var="run_id", filter_value =  "SAMPLE_CONTROL", additionnal_var=c(parameters$run_var, "assay_label", "block"))

  cv_data %>% group_by(!!sym(parameters$run_var), assay_label,block) %>% summarise(Mean_intraCV=mean(intra_CV, na.rm = TRUE),
                                                                             Median_intraCV=median(intra_CV,na.rm=TRUE),
                                                                             Mean_interCV=mean(inter_plate_CV, na.rm = TRUE),
                                                                             Median_interCV=median(inter_plate_CV,na.rm=TRUE), .groups = "drop") %>%  mutate(across(where(is.numeric), ~ round(., digits = 2))) %>% 
kbl(position = "!h",format="latex", booktabs = T, caption = "Mean and Median of CV based on samples controls per block") %>% kable_styling(font_size = 12,full_width = FALSE)

```




\newpage
\section{Distribution of Coefficients of variance based on samples controls}




```{r, results='asis', echo=FALSE}
  cat("![Distribution of intraCV based on samples controls](", output_dir, "/violin_intra.", plot_settings$format, "){width=100%}",sep="")

```

```{r, results='asis', echo=FALSE}
  cat("![Distribution of interCV based on samples controls](", output_dir, "/violin_inter.", plot_settings$format, "){width=100%}",sep="")

```



\newpage
\section{Coefficient of variance based on Healthy pool concordance samples}


```{r}

cv_data<-aggregate_cv(yourdata, "NPX", parameters$Background, comp_var="run_id", filter_col = "sample_info",  filter_value =  "Healthy pool", additionnal_var=c(parameters$run_var, "assay_label"))

  cv_data %>% group_by(!!sym(parameters$run_var), assay_label) %>% summarise(Mean_intraCV=mean(intra_CV, na.rm = TRUE),
                                                                             Median_intraCV=median(intra_CV,na.rm=TRUE),
                                                                             Mean_interCV=mean(inter_plate_CV, na.rm = TRUE),
                                                                             Median_interCV=median(inter_plate_CV,na.rm=TRUE), .groups = "drop") %>% 
mutate(across(where(is.numeric), ~ round(., digits = 2))) %>%     
kbl(position = "!h",format="latex", booktabs = T, caption = "Mean and Median of CV based on samples controls") %>% kable_styling(font_size = 12,full_width = FALSE)


```



\newpage
\section{Average detectability, based on concordance samples}


```{r}

detectability_data %>% group_by(!!sym(parameters$run_var)) %>% summarise(Average_detectability=round(mean(Detectability, na.rm = TRUE),digits = 3)) %>% 
kbl(position = "!h",format="latex", booktabs = T, caption = "Average detectability per Product") %>% kable_styling(font_size = 12,full_width = FALSE)

```


```{r}

detectability_data %>% group_by(!!sym(parameters$run_var),block) %>% summarise(Average_detectability=round(mean(Detectability, na.rm = TRUE),digits = 3)) %>% pivot_wider(names_from =!!sym(parameters$run_var), values_from = Average_detectability ) %>% 
kbl(position = "!h",format="latex", booktabs = T, caption = "Average detectability per Product and block") %>% kable_styling(font_size = 12,full_width = FALSE)

```


\newpage
\section{Across block distribution of assays loosing more than 10 percent detectability}


```{r}

 unique_categories <- unique(detectability_data[[parameters$run_var]])
  
  if (length(unique_categories) != 2) {
    stop("The comparison variable must have exactly two unique categories.")
  }
  
  cat1 <- unique_categories[1]
  cat2 <- unique_categories[2]
  
  
detectability_data %>%
    group_by(!!sym(parameters$run_var),block, olink_id) %>%
    summarise(agg_value = mean(Detectability, na.rm = TRUE), .groups = "drop") %>% # This in case we have more than one value for each comp var and olink id.
    pivot_wider(names_from = !!sym(parameters$run_var), values_from = agg_value) %>%
    group_by(block,olink_id) %>%
    summarise(diff_value = .data[[cat1]] - .data[[cat2]], .groups = "drop") %>% filter(diff_value > 0.10) %>% group_by(block) %>% count() %>% 
kbl(position = "!h",format="latex", booktabs = T, caption = "Assays loosing more than 10 percent detectability per block") %>% kable_styling(font_size = 12,full_width = FALSE)

```




\newpage
\section{Percentage of assays detectable in at least 10, 50 percent of the concordance samples}


```{r, echo=FALSE, results='asis'}

calculate_detectability_percentage_per_group(
     df = detectability_data, 
     assay_group = unique(yourdata %>% select(olink_id,assay_label)), 
     detectability_col = "Detectability", 
 thresholds = c( 0.1, 0.5),
   group_var = "assay_label",
   instrument_col = "Product") %>% kbl(position = "!h", format="latex", caption = "Percentage of assays detectable in at least 10, 50 percent of the concordance samples" ) %>%
        kable_classic(full_width = FALSE)

```


\newpage
\section{Overlap of assays detectable in at least 10, 50 percent of the concordance samples between products}


```{r, results='asis', echo=FALSE}
  cat("![Overlap of assays detectable in at least 10 percent of the concordance samples between products](", output_dir, "/Venn_detect.", plot_settings$format, "){width=100%}",sep="")

```

```{r}
rm("Concordance_count_npx_above_background.rds")
rm("Concordance_count_npx.rds")

```


\newpage
\section{Appendix}


\subsection{R Session Information}

```{r, results='asis'}
toLatex(sessionInfo())
```

